Parameter precision in economic growth models inspired by Ecological Economics

### 2015-10-04 Articles/Article_EcolEcon_2015/Paper0/Paper0-Resampling.Rnw
Editing intro.

\end{enumerate}
%

@@

Information about the C-D equation here.

%------------------------------
\subsubsection{CES production function}
\label{sec:CES}
%------------------------------

@@

The energy-augmented 
Constant Elasticity of Substitution (CES) production function%
  \footnote{
  We use the The \texttt{R} 
  % \citep{R} 
  package \texttt{micEconCES} 
  % \citep{Henningsen:2011td} 
  to estimate the CES production function.
  }
can be given as
%
\begin{equation} \label{eq:CESgeneric}
  y = \gamma \: A \: \left\{\delta \left[\delta_1 x_1^{-\rho_1} 
      + (1-\delta_1)x_2^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) x_3^{-\rho} \right\}^{-1/\rho}; 
      A \equiv \mathrm{e}^{\lambda (t-t_0)},
\end{equation}
%
where
$y \equiv Y/Y_{0}$,
$\gamma$ is a scale parameter,
e is the base of the natural logarithm, 
$\lambda$ is the Solow residual (often thougth to represent the pace of technological progress),
$t$ (time) is measured in years,
$x_1$, $x_2$, and $x_3$ are factors of production
and permutations of capital~($k$), labor~($l$), and energy~($e$),
$k \equiv K/K_{0}$, 
$l \equiv L/L_{0}$, 
$Y$ (economic output) is represented by GDP, 
$K$ (capital) is expressed in currency units, 
$L$ (labor) is expressed in workers or work-hours/year, and
the 0 subscript indicates values at an initial year.%
  \footnote{Dimensionless, indexed quantities are represented by 
  lower-case symbols 
  ($y$, $k$, $l$, $e$, $q$, $x$, and $u$), and dimensional 
  quantities are represented by upper-case symbols 
  ($Y$, $K$, $L$, $E$, $Q$, $X$, and $U$). 
  Model parameters are represented by Greek letters
  ($\alpha_1$, $\alpha_2$, $\lambda$, $\theta$).
  }
The CES model without energy is given in Equation~\ref{eq:CESkl}.
%
\begin{equation} \label{eq:CESkl}
  y = \gamma \: A \: \left[\delta_1 k^{-\rho_1} 
      + (1-\delta_1)l^{-\rho_1} \right]^{-1/\rho_1}; 
      A \equiv \mathrm{e}^{\lambda (t-t_0)} . 
\end{equation}
%
Equation~\ref{eq:CESkle} augments Equation~\ref{eq:CESkl} with energy 
using a ($kl$)($e$) nesting structure, as is typical in the literature. 
Equation~\ref{eq:CESkl} is a degenerate form of Equation~\ref{eq:CESkle} 
where $\delta \rightarrow 1$. 
%
\begin{equation} \label{eq:CESkle}
  y = \gamma \: A \: \left\{\delta \left[\delta_1 k^{-\rho_1} 
      + (1-\delta_1)l^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) e^{-\rho} \right\}^{-1/\rho}; 
      A \equiv \mathrm{e}^{\lambda (t-t_0)}.
\end{equation}
%
In the CES production function, 
$\gamma$ is a fitting parameter that accounts for an atypical first year. 
The fitting parameters $\rho_1$ and $\rho$ indicate elasticities 
of substitution ($\sigma_1$ and $\sigma$).
The elasticity of substitution 
between capital ($k$) and labor ($l$) is given by 
$\sigma_1 = \frac{1}{1+\rho_1}$, and
the elasticity of substitution between ($kl$) and ($e$) is given by 
$\sigma = \frac{1}{1+\rho}$. 
As $\rho_1 \rightarrow 0$, $\sigma_1 \rightarrow 1$,
and the embedded CES production function for $k$ and $l$ degenerates 
to the Cobb-Douglas production function.
Similarly, as $\rho \rightarrow 0$, $\sigma \rightarrow 1$,
and the CES production function for $(kl)$ and $(e)$ degenerates 
to the Cobb-Douglas production function.
As $\sigma \rightarrow \infty$ ($\rho \rightarrow -1$), 
($kl$) and ($e$) are perfect substitutes. 
As $\sigma \rightarrow 0$ ($\rho \rightarrow \infty$), 
($kl$) and ($e$) are perfect complements: 
no substitution is possible. Similarly, 
as $\sigma_1 \rightarrow 0$ ($\rho_1 \rightarrow \infty$),
$k$ and $l$ are perfect complements.
$\delta_1$ describes the relative importance of capital ($k$)
and labor ($l$), and
$\delta$ describes the importance of ($kl$) relative to ($e$).

The CES production function shown in Equation~\ref{eq:CESkl} 
includes two factors of production, capital~($k$) and labor~($l$).
Extending to three factors of production can be accomplished by nesting
two factors of production against the third.
\authNote{rjp:  
This is not really the correct metaphor for ``nest".  They are nested 
because one CES formula is inside the other (like nested tables).  Basically
the inner CES formula is being plugged into the slot of one of the factors of production
of the outer CES formula.}%
Equation~\ref{eq:CESkle} augments Equation~\ref{eq:CESkl} with energy~($e$),
in a ($kl$)$e$ nesting structure, as is common in the literature:
%
\begin{equation} \label{eq:CESkle}
  y = \theta \: A \: \left\{\delta \left[\delta_1 k^{-\rho_1}
      + (1-\delta_1)l^{-\rho_1} \right]^{\rho/\rho_1}
      + (1-\delta) e^{-\rho} \right\}^{-1/\rho} \; ; \;
      A \equiv \mathrm{e}^{\lambda t} \; .
\end{equation}
%
Note that Equation~\ref{eq:CESkl} is a degenerate form of Equation~\ref{eq:CESkle}
with $\delta = 1$ and $\rho$ and $e$ undetermined.
Two other nestings of the 
factors of production---($le$)$k$ and ($ek$)$l$---are possible 
with a three-factor CES function.
\authNoted{Do we need/want to comment about the other 3 nestings and that
each is just a reparameterization of one of the three we have given. ---rjp
I don't think we need to comment on the other three. ---MKH}
%
\begin{equation} \label{eq:CESlek}
  y = \theta \: A \: \left\{\delta \left[\delta_1 l^{-\rho_1}
      + (1-\delta_1) e^{-\rho_1} \right]^{\rho/\rho_1}
      + (1-\delta) k^{-\rho} \right\}^{-1/\rho};
      A \equiv \mathrm{e}^{\lambda t}
\end{equation}
%
\begin{equation} \label{eq:CESekl}
  y = \theta \: A \: \left\{\delta \left[\delta_1 e^{-\rho_1}
      + (1-\delta_1) k^{-\rho_1} \right]^{\rho/\rho_1}
      + (1-\delta) l^{-\rho} \right\}^{-1/\rho};
      A \equiv \mathrm{e}^{\lambda t}
\end{equation}
%
Each of the CES functions (Equations~\ref{eq:CESkl} 
and~\ref{eq:CESkle}--\ref{eq:CESekl}) assumes constant returns 
to scale.

Output elasticities~($\alpha$) are not constant;
rather, they vary with factors of production $k$, $l$, and $e$ over time.
For the $(kl)e$ nesting (Equation~\ref{eq:CESkle}),
output elasticities are shown in Equations~\ref{eq:alpha_k}--\ref{eq:alpha_e}.
Output elasticities for other nests can obtained by permutation of
the factors of production $k$, $l$, and $e$.
%
\begin{equation} \label{eq:alpha_k}
  \alpha_k \equiv
  \frac{
  \frac{\partial y}{\partial k} }{
  \frac{y}{k}
  } = \frac{\delta  \delta _1 k^{-\rho _1} \left[\delta _1 k^{-\rho _1}
  	+\left(1 - \delta_1 \right) 
		l^{-\rho _1}\right]{}^{\frac{\rho }{\rho _1}-1}}{\delta  
		\left[\delta _1
     k^{-\rho _1} + \left(1 - \delta _1 \right) l^{-\rho _1} \right]{}^{\frac{\rho }{\rho
     _1}} + (1 - \delta) e^{-\rho }}
\end{equation}
%
\begin{equation} \label{eq:alpha_l}
  \alpha_l \equiv 
  \frac{
  \frac{\partial y}{\partial l} }{
  \frac{y}{l}} =
  \frac{\delta  \left(1-\delta _1\right) l^{-\rho _1}
     \left[\delta _1 k^{-\rho _1}+\left(1 - \delta_1\right) l^{-\rho_1}\right]{}
	 				^{\frac{\rho }{\rho _1}-1}}
		{\delta  \left[\delta _1 k^{-\rho_1} + \left(1 - \delta_1\right) 
					l^{-\rho_1}\right]{}^{\frac{\rho }{\rho _1}}+ (1 - \delta) e^{-\rho }}
\end{equation}
%
\begin{equation} \label{eq:alpha_e}
  \alpha_e \equiv
  \frac{
  \frac{\partial y}{\partial e} }{
  \frac{y}{e}
  } =  \frac{1 - \delta}{\delta  e^{\rho } \left[\delta _1 k^{-\rho_1} 
  			+ \left(1 - \delta_1 \right) l^{-\rho _1}\right]{}
				^{\frac{\rho }{\rho_1}} + 1 - \delta}
\end{equation}
%
It can be verified that
%
\begin{equation} \label{eq:sum_betas}
  \alpha_k + \alpha_l + \alpha_e = 1
\end{equation}
%
for all nests.

Elasticities of substitution~($\sigma$) are a function of $\rho_1$, $\rho$, and
nesting as shown in 
Table~\ref{tab:CES_sigma_equations}.
\authNote{Need to use the equations for time-varying $\sigma$ values instead of this table.}
Note that $\sigma~\in~[0,\infty)$. 
The limiting case $\sigma \to 0$ is the Leontief production function and 
indicates perfect complementarity.  
The limiting case 
$\sigma \to \infty$ is a linear production function and 
indicates infinite substitutability.
When $\sigma = 1$, the CES production function simplifies to the
Cobb-Douglas production function.
%
\begin{table} \caption{Equations for elasticities of substitution~($\sigma$)
for various CES nests.}
\label{tab:CES_sigma_equations}
  \begin{center}
    \begin{tabular}{c c c c c c c c}
      \toprule
      Nest       & Equation
	  			 & $\sigma_{kl}$             & $\sigma_{(kl)e}$
	  			 & $\sigma_{le}$             & $\sigma_{(le)k}$
				 & $\sigma_{ek}$             & $\sigma_{(ek)l}$       \\
      \midrule
      ($kl$) \;  & \ref{eq:CESkl}
	  			 & $\frac{1}{1 + \rho_1}$   & --
	             & --                       & --
				 & --                       & --                    \\
      ($kl$)$e$  & \ref{eq:CESkle}
	  			 & $\frac{1}{1 + \rho_1}$   & $\frac{1}{1 + \rho}$
	             & --                       & --
				 & --                       & --                    \\
      ($le$)$k$  & \ref{eq:CESlek}
	  			 & --                       & --
	             & $\frac{1}{1 + \rho_1}$   & $\frac{1}{1 + \rho}$
				 & --                       & --                    \\
      ($ek$)$l$  & \ref{eq:CESekl}
	  			 & --                       & --
	             & --                       & --
	             & $\frac{1}{1 + \rho_1}$   & $\frac{1}{1 + \rho}$   \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

>>>>

\end{enumerate}
%
for the Portuguese economy from 1960--2010.

@@

**** Go completely generic everywhere ($x_1$ instead of $k$). ****

The Cobb-Douglas production function 
is the canonical representation of indexed economic output ($y$) 
as a function of 
indexed labor~($l$), 
indexed capital stock~($k$), and 
an additional exogenous, time-dependent term~($A$).
This production function can be expressed as
%
\begin{equation} \label{eq:CD}
  y = \theta A k^\alpha_k l^\alpha_l \; ; \; A \equiv \mathrm{e}^{\lambda(t-t_0)},
\end{equation}
%
where 
$y \equiv Y/Y_{0}$,
$\theta$ is a scale parameter,
e is the base of the natural logarithm, 
$\lambda$ is represents the pace of technological progress,
$t$ is time (typically measured in years),
$k \equiv K/K_{0}$, 
$l \equiv L/L_{0}$, 
$Y$ is represented by GDP, 
$K$ is expressed in currency units, 
$L$ is expressed in workers or work-hours/year, and
0 subscripts indicate values at an initial year.
The parameters $\alpha_k$ and $\alpha_l$ are output elasticities
that indicate the marginal productivity of 
capital and labor in the economy, respectively.
Typically, the constraint $\alpha_k + \alpha_l = 1$ is assumed, 
indicating constant returns to scale.%
  \footnote{Dimensionless, indexed quantities are represented by lower-case symbols 
  ($y$, $k$, $l$, and later $e$), 
  and dimensional quantities are represented by upper-case symbols 
  ($Y$, $K$, $L$, and later $E$). 
  Model parameters are represented by Greek letters
  ($\theta$, $\lambda$, $\alpha_k$, $\alpha_l$, 
  later $\alpha_e$, $\delta_1$, $\delta$, $\sigma_1$, and $\sigma$).}

The capital-labor Cobb-Douglas production function shown in Equation~\ref{eq:CD}
can be augmented by energy:
%
\begin{equation} \label{eq:CDe}
  y = \theta A k^\alpha_k l^\alpha_l e^\alpha_e \; ; \; A \equiv \mathrm{e}^{\lambda(t-t_0)},
\end{equation}
%
where $e \equiv E/E_0$, and $E$ is in units of energy per time, typically TJ/year. 
The energy-augmented Cobb-Douglas production function 
is often assumed to have constant returns to scale for the three factors 
of production: $\alpha_k + \alpha_l + \alpha_e = 1$.

@@

The Constant Elasticity of Substitution~(CES) model 
without energy is given in Equation~\ref{eq:CESkl}.
%
\begin{equation} \label{eq:CESkl}
  y = \gamma \: A \: \left[\delta_1 k^{-\rho_1} 
      + (1-\delta_1)l^{-\rho_1} \right]^{-1/\rho_1}; \;
      A \equiv \mathrm{e}^{\lambda (t-t_0)} . 
\end{equation}
%
The generic CES production function
can be given as
%
\begin{equation} \label{eq:CESgeneric}
  y = \gamma \: A \: \left\{\delta \left[\delta_1 x_1^{-\rho_1} 
      + (1-\delta_1)x_2^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) x_3^{-\rho} \right\}^{-1/\rho}; 
      A \equiv \mathrm{e}^{\lambda (t-t_0)},
\end{equation}
%
where
$x_1$, $x_2$, and $x_3$ are factors of production
and permutations of capital~($k$), labor~($l$), and energy~($e$).
%
Equation~\ref{eq:CESkle} augments Equation~\ref{eq:CESkl} with energy 
using a ($kl$)$e$ nesting structure, as is typical in the literature. 
Equation~\ref{eq:CESkl} is a degenerate form of Equation~\ref{eq:CESkle} 
where $\delta \rightarrow 1$. 
%
\begin{equation} \label{eq:CESkle}
  y = \gamma \: A \: \left\{\delta \left[\delta_1 k^{-\rho_1} 
      + (1-\delta_1)l^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) e^{-\rho} \right\}^{-1/\rho}; 
      A \equiv \mathrm{e}^{\lambda (t-t_0)}.
\end{equation}
%
In the CES production function, 
$\theta$ is a fitting parameter that accounts for an atypical first year. 
The fitting parameters $\rho_1$ and $\rho$ indicate elasticities 
of substitution ($\sigma_1$ and $\sigma$).

Two other nestings of the 
factors of production---($le$)$k$ and ($ek$)$l$---are possible 
with a three-factor CES function.
%
\begin{equation} \label{eq:CESlek}
  y = \theta \: A \: \left\{\delta \left[\delta_1 l^{-\rho_1}
      + (1-\delta_1) e^{-\rho_1} \right]^{\rho/\rho_1}
      + (1-\delta) k^{-\rho} \right\}^{-1/\rho};
      A \equiv \mathrm{e}^{\lambda t}
\end{equation}
%
\begin{equation} \label{eq:CESekl}
  y = \theta \: A \: \left\{\delta \left[\delta_1 e^{-\rho_1}
      + (1-\delta_1) k^{-\rho_1} \right]^{\rho/\rho_1}
      + (1-\delta) l^{-\rho} \right\}^{-1/\rho};
      A \equiv \mathrm{e}^{\lambda t}
\end{equation}
%
Each of the CES functions (Equations~\ref{eq:CESkl} 
and~\ref{eq:CESkle}--\ref{eq:CESekl}) assumes constant returns 
to scale.

For the CES function, 
output elasticities~($\alpha$) are not constant;
rather, they vary with factors of production $k$, $l$, and $e$ over time.
For the $(kl)e$ nesting (Equation~\ref{eq:CESkle}),
output elasticities are shown in Equations~\ref{eq:alpha_k}--\ref{eq:alpha_e}.
Output elasticities for other nests can obtained by permutation of
the factors of production $k$, $l$, and $e$.
%
\begin{equation} \label{eq:alpha_k}
  \alpha_k \equiv
  \frac{
  \frac{\partial y}{\partial k} }{
  \frac{y}{k}
  } = \frac{\delta  \delta _1 k^{-\rho _1} \left[\delta _1 k^{-\rho _1}
  	+\left(1 - \delta_1 \right) 
		l^{-\rho _1}\right]{}^{\frac{\rho }{\rho _1}-1}}{\delta  
		\left[\delta _1
     k^{-\rho _1} + \left(1 - \delta _1 \right) l^{-\rho _1} \right]{}^{\frac{\rho }{\rho
     _1}} + (1 - \delta) e^{-\rho }}
\end{equation}
%
\begin{equation} \label{eq:alpha_l}
  \alpha_l \equiv 
  \frac{
  \frac{\partial y}{\partial l} }{
  \frac{y}{l}} =
  \frac{\delta  \left(1-\delta _1\right) l^{-\rho _1}
     \left[\delta _1 k^{-\rho _1}+\left(1 - \delta_1\right) l^{-\rho_1}\right]{}
	 				^{\frac{\rho }{\rho _1}-1}}
		{\delta  \left[\delta _1 k^{-\rho_1} + \left(1 - \delta_1\right) 
					l^{-\rho_1}\right]{}^{\frac{\rho }{\rho _1}}+ (1 - \delta) e^{-\rho }}
\end{equation}
%
\begin{equation} \label{eq:alpha_e}
  \alpha_e \equiv
  \frac{
  \frac{\partial y}{\partial e} }{
  \frac{y}{e}
  } =  \frac{1 - \delta}{\delta  e^{\rho } \left[\delta _1 k^{-\rho_1} 
  			+ \left(1 - \delta_1 \right) l^{-\rho _1}\right]{}
				^{\frac{\rho }{\rho_1}} + 1 - \delta}
\end{equation}
%
It can be verified that
%
\begin{equation} \label{eq:sum_betas}
  \alpha_k + \alpha_l + \alpha_e = 1
\end{equation}
%
for all nests.

Elasticities of substitution~($\sigma$) between factors of production
are a function of $\rho_1$, $\rho$, and nesting. 
**** More here on the actual $\sigma$ equations. 
Gather them from the Henningsen and Henningsen paper. ****

@@

Teaser at the end that variability varies from one country to the next.
